{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2025-10-23 lagrangian.tex
Small changes to R-parity

%%%%%%%%%%%%%%%%
\section{$R$-symmetry}
\label{sec:R_symmetry}
%%%%%%%%%%%%%%%%
The Lagrangian we have constructed is invariant under both internal gauge transformations and external (coordinate) super-Poincaré transformations, where the latter has been built in through the superspace coordinates. The Minkowski part of superspace, the four-vector coordinates, transform under the Lorentz group, and since all our Lagrangian ingredients are Lorentz scalars they are invariant. We can now ask the question if there is any transformation of the superspace coordinates $\theta$ and $\bar\theta$ that the Lagrangian is also invariant under.

In Sec.~\ref{sec:superalgebra} we  saw that the only possible nontrivial interaction between a gauge group and the supersymmetry generators for $N=1$ supersymmetry is given by  (\ref{eq:QR_commutator_dirac}), which in Weyl spinor notation for the supercharges is
\begin{equation}
[Q_A, R] = Q_A, \quad  [\bar Q_{\dot A}, R ] = -\bar Q_{\dot A},
\end{equation}
where $R$ is the generator for an abelian $U(1)_R$ gauge group. If we impose such an  {\bf $\mathbf R$-symmetry}  it must act differently on the different components of a superfield since it does not commute with the supercharges.

@@

\begin{figure}
\centering
\begin{tikzpicture}
\begin{feynman}
%
\diagram [xshift=6cm, yshift=-0.6cm, layered layout, horizontal=b to c] {
a [particle=$h$] -- [scalar] b [dot, black,  label=$\lambda_s$]
-- [scalar] c [particle=$h$],
};
\path (b)--++(90:0.8) coordinate (A);
\path (b)--++(90:1.6) coordinate (B);
\draw [dashed] (A) circle(0.8);
\draw (B) node[above] {$s$};
%
\diagram [layered layout, horizontal=b to c] {
a [particle=$h$] -- [scalar] b [dot, black, label=$\lambda_f\hphantom{m}$]
-- [fermion, half left, out=90, in=90, edge label = $f$] c [dot, black, label=$\hphantom{nm}\lambda_f$]
-- [fermion, half left, out=90, in=90, edge label = $\bar{f}$] b,
c -- [scalar] d [particle=$h$],
};
%
\diagram [xshift=10cm, layered layout, horizontal=b to c] {
a [particle=$h$] -- [scalar] b [dot, black, label=$\lambda_V\hphantom{m}$]
-- [boson, half left, out=90, in=90, edge label = $V$] c [dot, black, label=$\hphantom{nm}\lambda_V$]
-- [boson, half left, out=90, in=90, edge label = $V$] b,
c -- [scalar] d [particle=$h$],
};
%
\end{feynman}
\end{tikzpicture}
\caption{Higgs self-energy diagrams with  scalars $s$ (left),  fermions $f$ (middle), and vectors $V$ (right). The dots indicate vertices with the given coupling strength.}  
\label{fig:hierarchy}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%
\section{$R$-symmetry}
\label{sec:R_symmetry}
%%%%%%%%%%%%%%%%
The Lagrangian we have constructed is invariant under both internal gauge transformations and external (coordinate) super-Poincaré transformations, where the latter has been built in through the superspace coordinates. In Sec.~\ref{sec:superalgebra} we  saw that the only possible nontrivial interaction between a gauge group and the supersymmetry generators for $N=1$ supersymmetry is given by  (\ref{eq:QR_commutator_dirac}), which in Weyl spinor notation for the supercharges is
\begin{equation}
[Q_A, R] = Q_A, \quad  [\bar Q_{\dot A}, R ] = -\bar Q_{\dot A},
\end{equation}
where $R$ is the generator for an abelian $U(1)_R$ gauge group. If we impose such an  {\bf $\mathbf R$-symmetry}  it must act differently on the different components of a superfield since it does not commute with the supercharges.

@@

\begin{figure}
\centering
\begin{tikzpicture}
\begin{feynman}
%
\diagram [yshift=-0.6cm, layered layout, horizontal=b to c] {
a [particle=$h$] -- [scalar] b [dot, black,  label=$\lambda_s$]
-- [scalar] c [particle=$h$],
};
\path (b)--++(90:0.8) coordinate (A);
\path (b)--++(90:1.6) coordinate (B);
\draw [dashed] (A) circle(0.8);
\draw (B) node[above] {$s$};
%
\diagram [xshift=5cm, layered layout, horizontal=b to c] {
a [particle=$h$] -- [scalar] b [dot, black, label=$\lambda_f\hphantom{m}$]
-- [fermion, half left, out=90, in=90, edge label = $f$] c [dot, black, label=$\hphantom{nm}\lambda_f$]
-- [fermion, half left, out=90, in=90, edge label = $\bar{f}$] b,
c -- [scalar] d [particle=$h$],
};
%
\diagram [xshift=10cm, layered layout, horizontal=b to c] {
a [particle=$h$] -- [scalar] b [dot, black, label=$\lambda_V\hphantom{m}$]
-- [boson, half left, out=90, in=90, edge label = $V$] c [dot, black, label=$\hphantom{nm}\lambda_V$]
-- [boson, half left, out=90, in=90, edge label = $V$] b,
c -- [scalar] d [particle=$h$],
};
%
\end{feynman}
\end{tikzpicture}
\caption{Higgs self-energy diagrams with  scalars $s$ (left),  fermions $f$ (middle), and vectors $V$ (right). The dots indicate vertices with the given coupling strength.}  
\label{fig:hierarchy}
\end{figure}
